---
title: 'ESM 244 Lab 3 Part 1: Binary Logistic Regression'
author: "Minnie Ringland"
date: "1/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(GGally)
library(broom)
library(jtools)
```

```{r}
# Explore the penguins dataset
#penguins %>% 
#  select(species, bill_length_mm:sex) %>% 
#  ggpairs(aes(color=species))
```

Create a subset with only Adelie and Chinstrap penguins
```{r}

#levels(penguins$species)

adelie_chinstrap <- penguins %>%
  filter(species %in% c("Adelie","Chinstrap")) %>% 
  mutate(species = fct_drop(species)) # drop any factor levels for which there are no observations left

#levels(adelie_chinstrap$species)

```

Let's just dataviz of those two species:
```{r}
ggplot(data=adelie_chinstrap, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color=sex)) +
  facet_wrap(~species)

```
We see some sexual dimorphism (males>females) in both physical traits. Looks like maybe Chinstraps tend to be bigger in flipper length and mayyybe larger body mass.

Once we've adequately explored the data, let's use binary logistic regression to see how these variables can predict penguin species:
- body mass
- flipper length
- penguin sex

```{r}
adelie_chinstrap_blr <- glm(species~ body_mass_g + flipper_length_mm + sex,
                            data = adelie_chinstrap,
                            family = "binomial")

adelie_chinstrap_blr

summary(adelie_chinstrap_blr)
```

```{r}
ggplot(data=adelie_chinstrap, aes(x=species, y = flipper_length_mm)) +
  geom_jitter(aes(color=sex))
```

What are the actual probabilities of being Chinstrap for each of the existing observations in our adelie_chinstrap dataframe?

```{r}
blr_fitted <- adelie_chinstrap_blr %>% 
  broom::augment(type.predict="response")
```

Plotting fitted probabilities:
```{r}
ggplot(data=blr_fitted, aes(x=flipper_length_mm, y = .fitted)) +
  geom_point(aes(color=sex, pch=species)) +
  geom_smooth(aes(color=sex), se = FALSE) +
  labs(x="Flipper length(mm)",
       y="Probability of outcome Chinstrap")
```

Plot model based on different variables:
```{r}

effect_plot(adelie_chinstrap_blr,
            pred = flipper_length_mm,
            y.label = "Probability of Chinstrap",
            interval = TRUE)

effect_plot(adelie_chinstrap_blr,
            pred = body_mass_g,
            y.label = "Probability of Chinstrap",
            interval = TRUE)

```


What's the probability that a female penguin weighing 3410g with a flipper length of 192mm is a Chinstrap?

```{r}
ex_1 <- predict(adelie_chinstrap_blr,
                data.frame(sex="female",
                           body_mass_g = 3410,
                           flipper_length_mm = 192),
                type = "response") # without this, gives log odds

ex_1
```

